\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-08-02 NAR_DRAFT.tex
changes

\subsection{Estimation of the mutation rate.}
The final piece of equation \ref{eqn:detailed_prior} to specify is $p(M)$, the prior probability that a mutation will occur at a given site, which usually specified as the mutation \textit{rate} per base $\mu$.
In an exponentially growing tumor, the number of new mutations per cell division is the product of the number of bases $N$, the per base mutation rate $\mu$, and the \textit{effective} birth rate $\beta$, which is a function of the birth $b$ and death $d$ rates.
\citet{Williams2016} showed that if 
$$
  \beta = \frac{b-d}{b},
$$
then the number of new mutations per cell division is
$$
  \frac{N \mu}{\beta}.
$$
We make the simplifying assumption that $d = 0$, such that $\beta = 1$
In a neutrally or nearly neutrally evolving tumor, they further show that the total number of mutations $\mathrm{M}$ between two allele frequencies ($f_{min}$,$f_{max}$) is
\begin{equation}
  \label{eqn:mut_rate}
$$
  \mathrm{M}(f_{min},f_{max}) = N\frac{\mu}{\beta}\left(\frac{1}{f_{min}} - \frac{1}{f_{max}}\right).
$$
\end{equation}

>>>>

\subsection{Estimation of the mutation rate.}
The final piece of equation \ref{eqn:detailed_prior} to specify is $p(M)$, the prior probability that a mutation will occur at a given site, which usually specified as the mutation \textit{rate} per base $\mu$.
In an exponentially growing tumor, the number of new mutations per cell division is the product of the number of bases $N$, the per base mutation rate $\mu$, and the \textit{effective} birth rate $\beta$, which is a function of the birth $b$ and death $d$ rates.
\citet{Williams2016} showed that if 
$$
  \beta = \frac{b-d}{b},
$$
then the number of new mutations per cell division is
$$
  \frac{N \mu}{\beta}.
$$
In a neutrally or nearly neutrally evolving tumor, they further show that the total number of mutations $\mathrm{M}$ between two allele frequencies ($f_{min}$,$f_{max}$) is
\begin{equation}
  \label{eqn:mut_rate}
$$
  \mathrm{M}(f_{min},f_{max}) = N\frac{\mu}{\beta}\left(\frac{1}{f_{min}} - \frac{1}{f_{max}}\right).
$$
\end{equation}
